%%____________________________________________________________________ 
%% File: ReleaseNotes.tex
%%____________________________________________________________________ 
%%  
%% Author: Shaun ASHBY <Shaun.Ashby@cern.ch>
%% Update: 2005-11-02 17:09:41+0100
%% Revision: $Id: ReleaseNotes.tex,v 1.6 2007/02/27 16:16:01 sashby Exp $ 
%%
%% Copyright: 2005 (C) Shaun ASHBY
%%
%%--------------------------------------------------------------------
\chapter{SCRAM Release Notes}\label{ch:releasenotes}
\index{SCRAM!release notes}

\section{Version \thisrelease}
\begin{itemize}
\item Full support for \texttt{XML} in all \scram\ documents. Older
  document markup is no longer supported (it is therefore a requirement
  that all documents are migrated to proper \texttt{XML} syntax before
  this release can be used).

\item Added support for new types of build products: \texttt{plugin} and
  \texttt{unittest}. How these are handled by the Build System can be
  controlled via the build templates.

\item Bug fixes (\#XXXX, \ldots). Full details can be found in the bug
  digest.

\end{itemize}


\section{Version \lastrelease}
\begin{itemize}
\item Numerous bug fixes (\#14639, \#15207, \#15675, \#18267,\#18783,\#78421,\#19551). Full details
  can be found in the bug digest.
  
\end{itemize}

\section{Version V1\_0\_2}
\begin{itemize}
\item Numerous bug fixes (\#7570, \#7645, \#7731, \#8040,
  \#8067, \#8328, \#10306, \#10623, \#10774, \#11402). Full details
  can be found in the bug digest.

\item Support for XML in \buildfile s and full support for XML
  \buildfile s in the build system. Note that a command is
  provided to automatically migrate all \buildfile s from old 
  style to XML tag syntax. The build command also has the option to parse XML
  versions of \buildfile s. Note that, eventually, support for
  old (non-XML) syntax will be dropped.
  
\item Fix to install process to permit correct operation on 64-bit machines.
  
\item Updated \scram\ manual (\LaTeX) and doxygen reference manual
  (useful for reporting bugs or for people interested in
  browsing the code). Note that \texttt{doxygen} v1.4.0 (at least) should
  be used. 
  
\item Added \scram\ completion function for the \texttt{ZSH}
  environment. This allows automatic completion of current
  \scram\ commands and command arguments.
  
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%% The REAL End %%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "SCRAM-manual"
%%% End:

%%____________________________________________________________________ 
%% End of ReleaseNotes.tex
%%____________________________________________________________________ 
%%  
